Convex Surrogates via Polyhedral Losses

### 2019-02-02 cvx-fin.tex
intro pass

\begin{abstract}
  Convex surrogates are sweet.
  Given a loss for a classification-like problem, there are two natural approaches to design convex surrogates.
  First, one may attempt to map each prediction to a low-dimensional vector, and try to find a convex loss in that space with the right calibration.
  Second, one may simply try to find a surrogate within the class of piecewise-linear convex, or polyhedral, losses.
  We show an equivalence between these two approaches, and \raf{more stuff}.
  We show that every loss with a finite number of predictions has a convex surrogate in the above sense using one fewer dimension that the number of outcomes, and give a full characterization of the losses needing only $d$ dimensions for such a surrogate.
  We then apply this characterization to show novel lower bounds for abstain loss, demonstrating the power of our techniques over \raf{check this} alternatives such as feasible subspace dimension.
\end{abstract}
\begin{keywords}%
  information elicitation, proper scoring rules, surrogate loss functions
\end{keywords}

@@

Convex surrogate losses are a central building block in machine learning for classification and classification-like problems.
A growing body of research is concerned with designing and analyzing convex surrogates for specific problems, and more broadly, understanding when such surrogates can and cannot be found.\raf{cite}
In particular, recent work has developed tools to bound the number of dimensions required in order for there to exist a consistent convex surrogate loss.\raf{cite}
Yet in some cases these bounds are far from tight, such as for \emph{abstain loss} (classification with an abstain option)\raf{cite}, and many questions remain unanswered.

We augment this literature by studying a particularly natural approach for finding convex surrogates, wherein one \emph{embeds} a discrete loss function into a real vector space so that it can be convexified.
Special cases include hinge loss as a surrogate for 0-1 loss, and the abstain surrogate mentioned above.\raf{cite}
Using tools from property elicitation, we show a tight relationship between such embeddings and the class of polyhedral (i.e.\ piecewise-linear convex) loss functions: whenever such an embedding is possible, the discrete loss can be convexified into a polyhedral loss, and moreover, every polyhedral loss function is the result of such an embedding of some discrete loss.

We go on to study the \emph{embedding dimension}: the minimal dimension for a given loss to be so embedded.
Our results give a complete characterization in the case of dimension 1 (real-valued surrogates):
a discrete loss can be embedded in the real line if and only if it is order-sensitive, meaning there is an ordering of the predictions such that the loss increases as one moves away from the correct prediction.
We show further that order-sensitivity characterizes a much broader condition: the existence of any real-valued convex surrogate.
In higher dimensions, we give a characterization of embeddability in terms of convex polytopes, which yields tight bounds for abstain loss.
We conclude with several conjectures and directions for future work.

>>>>

\begin{abstract}
  \bo{TODO - iterate!}
  Convex surrogates are sweet.
  Given a loss for a classification-like problem, there are two natural approaches to design convex surrogates.
  First, one may attempt to map each prediction to a low-dimensional vector, and try to find a convex loss in that space with the right calibration.
  Second, one may simply try to find a surrogate within the class of piecewise-linear convex, or polyhedral, losses.
  We show an equivalence between these two approaches, and \raf{more stuff}.
  We show that every loss with a finite number of predictions has a convex surrogate in the above sense using one fewer dimension that the number of outcomes, and give a full characterization of the losses needing only $d$ dimensions for such a surrogate.
  We then apply this characterization to show novel lower bounds for abstain loss, demonstrating the power of our techniques over \raf{check this} alternatives such as feasible subspace dimension.
\end{abstract}
\begin{keywords}%
  information elicitation, proper scoring rules, surrogate loss functions
\end{keywords}

@@

Convex surrogate losses are a central building block in machine learning for classification and classification-like problems.
A growing body of research is concerned with designing and analyzing convex surrogates for given loss functions, and more broadly, understanding when such surrogates can and cannot be found.\raf{cite}
For example, recent work has developed tools to bound the required number of dimensions of the surrogate's hypothesis space.\raf{cite}
Yet in some cases these bounds are far from tight, such as for \emph{abstain loss} (classification with an abstain option)\raf{cite}.
Furthermore, the kinds of strategies available for constructing surrogates, and their relative power, are not well-understood.

We augment this literature by studying a particularly natural approach for finding convex surrogates, wherein one \emph{embeds} a discrete loss function into a real vector space so that it can be convexified.
Special cases include hinge loss as a surrogate for 0-1 loss and the abstain surrogate mentioned above.\raf{cite}
\bo{it felt like the following results were only a warmup instead of main results. Not sure how to change this, maybe a paragraph break helps.}

Using tools from property elicitation, we show a tight relationship between such embeddings and the class of polyhedral (i.e.\ piecewise-linear convex) loss functions: whenever such an embedding is possible, the discrete loss can be convexified into a polyhedral loss, and moreover, every polyhedral loss function is the result of such an embedding of some discrete loss.

We go on to study the \emph{embedding dimension}: the minimal dimension for a given loss to be so embedded.
Our results give a complete characterization in the case of dimension 1 (real-valued surrogates):
a discrete loss can be embedded in the real line if and only if it is order-sensitive, meaning there is an ordering of the predictions such that the loss increases as one moves away from the correct prediction.
We show further that order-sensitivity characterizes a much broader condition: the existence of any real-valued convex surrogate.
In other words, in one dimension the embedding approach to surrogate construction loses no generality.
In higher dimensions, we give a characterization of embeddability in terms of convex polytopes, which yields tight bounds for abstain loss.
We conclude with several conjectures and directions for future work.
